---
title: "HW5"
author: "Giang Vu"
date: "2/22/2021"
output: pdf_document
---

### (c)

```{r echo=FALSE, warning=FALSE,out.width="80%"}
#libraries
library(MASS)

#import data
q1c <- data.frame(read.table(file = "/Users/giangvu/Desktop/STAT 2132 - Applied Stat Method 2/HW/HW5/Q1c.txt", header = T, sep = "\t", stringsAsFactors = F))
q1c$X <- as.numeric(q1c$X)
q1c <- q1c[1:200,]

# #fit model Mi
q1c$Y <- factor(as.numeric(q1c$Y), levels = c(1,2,3), labels = c("1","2","3"), ordered = T)
# q1c.fit <- MASS::polr(Y ~ X , data = q1c)


###### 
#Mi
q1c$Y1 <- ifelse(q1c$Y==1,1,0)
q1c$Y1 <- factor(as.numeric(q1c$Y1), levels = c(0,1), labels = c("0","1"), ordered = T)
q1c.fit1 <- glm(Y1 ~ X, family = binomial(link = "logit"), data = q1c)



#fit model Mii
q1c$Y2 <- ifelse(q1c$Y==3,0,1)
q1c$Y2 <- factor(as.numeric(q1c$Y2), levels = c(0,1), labels = c("0","1"), ordered = T)

q1c.fit2 <- glm(Y2 ~ X, family = binomial(link = "logit"), data = q1c)


```
The fitted models are  

#### (Mi) 
$logit(\hat{\pi}_{i1}) = 0.5932 + 1.9634x_i$

#### (Mii)
$logit(\hat{\pi}_{i1}+\hat{\pi}_{i2}) = 3.707 + 2.258x_i$

#### (i) & (ii)
Below is the plot of estimated probabilities from (Mi) and the estimated probabilities from (Mii) as a function of xi (red and blue), as well as the logit of them as a function of xi (orange and green).
```{r echo=FALSE, warning=FALSE,out.width="80%"}
#plot for (i) - red + blue
plot(x=q1c$X,y= q1c.fit1$fitted.values,col="red",ylim = c(0,1),ylab="Estimated Probabilities", xlab = "X")
points(x=q1c$X,y=q1c.fit2$fitted.values,col="blue")
legend(-3.8,0.9,pt.bg = c("red","blue"),pch = c(1,1),legend = c("Model 1", "Model 2"),col = c("red","blue"))
```

```{r echo=FALSE, warning=FALSE,out.width="80%"}
#plot for (ii) - orange +green
plot(x=q1c$X,y = log(q1c.fit1$fitted.values/(1-q1c.fit1$fitted.values)),col="orange",ylab="Logit of Estimated Probabilities", xlab = "X",ylim = c(-8,10))
points(x=q1c$X,y = log(q1c.fit2$fitted.values/(1-q1c.fit2$fitted.values)),col="green")
legend(-3.8,8,pt.bg = c("orange","green"),pch = c(1,1),legend = c("Model 1", "Model 2"),col = c("orange","green"))


```



### (d)
```{r echo=FALSE, warning=FALSE,out.width="80%"}
#libraries
library(MASS)

#import data
q1d <- data.frame(read.table(file = "/Users/giangvu/Desktop/STAT 2132 - Applied Stat Method 2/HW/HW5/Q1d.txt", header = T, sep = "\t", stringsAsFactors = F))
q1d$X <- as.numeric(q1d$X)

#fit model Mi & Mii
q1d$Y <- factor(as.numeric(q1d$Y), levels = c(1,2,3), labels = c("1","2","3"), ordered = T)
#q1d.fit <- MASS::polr(Y ~ X , data = q1d)

q1d$Y1 <- ifelse(q1d$Y==1,1,0)
q1d$Y1 <- factor(as.numeric(q1d$Y1), levels = c(0,1), labels = c("0","1"), ordered = T)
q1d.fit1 <- glm(Y1 ~ X, family = binomial(link = "logit"), data = q1d)

#fit model Mii
q1d$Y2 <- ifelse(q1d$Y==3,0,1)
q1d$Y2 <- factor(as.numeric(q1d$Y2), levels = c(0,1), labels = c("0","1"), ordered = T)

q1d.fit2 <- glm(Y2 ~ X, family = binomial(link = "logit"), data = q1d)

```
Using another sample, the fitted models are  

#### (Mi) 
$logit(\hat{\pi}_{i1}) = 1.584 + 4.684x_i$

#### (Mii)
$logit(\hat{\pi}_{i1}+\hat{\pi}_{i2}) = 2.441 + 1.153x_i$

#### (i) & (ii)
Below is the plot of estimated probabilities from (Mi) and the estimated probabilities from (Mii) as a function of xi (red and blue), as well as the logit of them as a function of xi (orange and green).

```{r echo=FALSE, warning=FALSE,out.width="80%"}
#plot for (i) - red + blue
plot(x=q1d$X,y= q1d.fit1$fitted.values,col="red",ylim = c(0,1),ylab="Estimated Probabilities", xlab = "X")
points(x=q1d$X,y=q1d.fit2$fitted.values,col="blue")
legend(-3.8,0.9,pt.bg = c("red","blue"),pch = c(1,1),legend = c("Model 1", "Model 2"),col = c("red","blue"))

```

```{r echo=FALSE, warning=FALSE,out.width="80%"}
#plot for (ii) - orange +green
plot(x=q1d$X,y=  log(q1d.fit1$fitted.values/(1-q1d.fit1$fitted.values)),col="orange",ylab="Logit of Estimated Probabilities", xlab = "X",ylim =c(-8,10) )
points(x=q1d$X,y=log(q1d.fit2$fitted.values/(1-q1d.fit2$fitted.values)),col="green")
legend(-3.8,8,pt.bg = c("orange","green"),pch = c(1,1),legend = c("Model 1", "Model 2"),col = c("orange","green"))

```

### (e)

From the plots in parts (c) and (d), I am more comfortable fitting the proportional odds model to Q1d.txt sample because in Q1d.txt data, the plots of two different models cross each other. In the proportional odds model given, we see that the two models we want to fit have the same slope $\beta$. Therefore we would expect the plots of two models to be kind of parallel in Q1c.txt, not crossing like Q1d.txt



