---
title: "hw2"
author: "Giang Vu"
date: "2/3/2021"
output: pdf_document
---

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
library(ggrepel)
library(ggplot2)
library(dplyr)
library(Hmisc)
library(readxl)
library(DBI)
library(odbc)
library(readr)
library(writexl)
library(knitr)
library(clipr)
library(car)
library(MASS)
library(np)
library(olsrr)
library(chemometrics)
library(pls)
library(glmnet)
library(bootstrap)
library(boot)
```

## Homework 2

### 1.

#### (a)

The starting points for Gauss-Newton I chose are 25 for gamma0 and 1 for gamma1.  
  From the equation  
  $1/E(Y_i) = (1/X_i)\gamma_1/\gamma_0 + 1/\gamma_0$  
  $\rightarrow$ As X approaches infinity, $1/X_i$ approaches 0.   
  $\rightarrow$ Then $1/E(Y_i)$ approaches $1/\gamma_0$, or $Y_i$ approaches $\gamma_0$.  
  $\rightarrow$ $\gamma_0$ is the horizontal asymptote of this function.   
  From the plot of the data, the points seem to converge towards near 25, so I picked 25 as the starting point.   
  Also, as X increases, Y also increases, so I expect $\gamma_1$ to be positive, and the magnitude of it tells us how fast Y increases as X increases, so I picked the starting value 1.   

```{r echo=TRUE, warning=FALSE}
#read data
enzyme <- read.delim("/Users/giangvu/Desktop/STAT 2132 - Applied Stat Method 2/HW/HW2/Enzyme.txt")

#plot data
plot(enzyme$X, enzyme$Y, xlab="Concentration", ylab="Velocity")
```

#### (b)

Gamma0 and gamma1 are estimated using the starting points from part (a) below
 
```{r echo=TRUE, warning=FALSE}
#nls
out.start <- nls(Y ~ (gamma0*X)/(gamma1 + X), data = enzyme, start = list(gamma0=25,gamma1=1))
#estimates for gamma's
gamma.hat <- out.start$m$getAllPars()
gamma.hat
```

### 2.

#### (a)

The fit seems to be adequate.

```{r echo=TRUE, warning=FALSE}
#plot data 
plot(enzyme$X, enzyme$Y, xlab="Concentration", ylab="Velocity")
#add the nonlinear regression function line
lines(enzyme$X, out.start$m$fitted(), col="red")
```

#### (b)

Based on the residual plot and the qq-plot, the assumptions of constant variances and normality are both satisfied. The residual plots show no trend, and the tail of the qq-plot shows no skewness.

```{r echo=TRUE, warning=FALSE}
#residual plot - constant variances - satisfied
plot(out.start$m$fitted(), out.start$m$resid(), xlab="Fitted values", ylab="Residuals"); abline(h=0, col="red", lty=2)
#qq plot - normality
qqnorm(out.start$m$resid(), ylab="Residuals")
qqline(out.start$m$resid(), col="red")
```

#### (c)

Null hypothesis: Gamma1 = 20  
  Alternative hypothesis: Gamma1 different from 20  
  Assumptions: The residuals are not correlated with X, and with each other. They are also normally distributed.
  The test statistic is -9.73141 with 16 degrees of freedom.  
  The p-value is basically 0, very small   
  The 95% CI is (11.23,13.91)  
  We reject the null hypothesis, we have strong evidence to conclude that gamma1 cannot be equal to 20.    
  
```{r echo=TRUE, warning=FALSE}
#test gamma1 = 20
#Jacobian
J <- out.start$m$gradient()   #The 44 x 2 gradient matrix. This acts as our design matrix.
#MSE
sigma2 <- sum(out.start$m$resid()^2)/(nrow(J)-ncol(J)) #nrow(J) = n, ncol(J) = p
#SE for inference
se.gamma0 <- sqrt(sigma2)*sqrt( solve(t(J)%*%J)[1,1] )
se.gamma1 <- sqrt(sigma2)*sqrt( solve(t(J)%*%J)[2,2] )
#t-stat for test gamma1 = 20, df = n-p = 16 
t.gam1 <- (gamma.hat[2]-20)/se.gamma1

#CI
CI.gamma0 <- gamma.hat[1] + c(-1,1)*se.gamma0*qt(p = 0.95, df = nrow(J)-ncol(J))
CI.gamma1 <- gamma.hat[2] + c(-1,1)*se.gamma1*qt(p = 0.95, df = nrow(J)-ncol(J))
```

### 3.

After running a bootstrap with 1000 samples, the 95% percentile confidence intervals for gamma1 obtained is (11.03, 13.94). This is quite close compared to the interval based on the large sample theory we obtained in problem 2.

```{r echo=TRUE, warning=FALSE}
#bootstrap with 1000 samples
set.seed(200)
enzyme_bootfcn <- function(x,i){
  d<-x[i,]
  out.boot <- nls(Y ~ (gamma0*X)/(gamma1 + X), data = d, start = list(gamma0=25,gamma1=1))
  gamma.boot <- out.boot$m$getAllPars()
  return(gamma.boot[2])
}
enzyme_bootobj <- boot(enzyme,enzyme_bootfcn,R=1000)
head(enzyme_bootobj$t)
#bootstrap 95% percentile CI
enzyme_bootci <- boot.ci(enzyme_bootobj)
enzyme_bootci
```

