---
title: "hw3"
author: "Giang Vu"
date: "2/10/2021"
output: pdf_document
---

## Homework 3

### (a)

The model I am assuming is  
  $logit(\pi_i)=\beta_0+\beta_1log_2(X_i)$  
  where $\pi_i=Y_i/m$ is the mortality rate.   
  $\beta_0$ is log odds of $Y_i=1$ when $log_2(X_i)=0$.  
  $\beta_1$ is the log odds ratio between log dose $X_{i+1}$ and log dose $X_i$.   
  $Y_i|X_i$ are assumed to be independent.   

### (b)

After fitting the model using quasibinomial family with logic link, the plot of the fitted mortality rate and raw mortality fractions y/m against log dose is obtained below. Looking at the fit of the model, we could see that it fits quite well. I checked residual plot and qq-plot and the constant variances and normality assumptions are both satisfied as well.

```{r echo=FALSE, warning=FALSE,out.width="80%"}
#create df
toxic <- data.frame(log2x = c(0,1,2,3,4,5),
                    mortality = c(0/7,2/9,3/8,5/7,7/9,10/11))

#fit logistic model #used quasi because Yi is not exactly binomial
toxic.fit <- glm(mortality ~ log2x, family = quasibinomial(link = "logit"), data = toxic)
toxic.sum <- summary(toxic.fit)

#plot raw data
plot(toxic$log2x, toxic$mortality, xlab="Log Dose", ylab="Mortality Rate")
#plot fitted
lines(toxic$log2x, fitted(toxic.fit), col="red")
#model assumed in part a seems reasonable

```

```{r echo=FALSE, warning=FALSE,out.width="80%"}
#residual plot - constant variances - satisfied
plot(fitted(toxic.fit), toxic.fit$residuals, xlab="Fitted values", ylab="Residuals",main = "Residual plot"); abline(h=0, col="red", lty=2)
```

```{r echo=FALSE, warning=FALSE,out.width="80%"}
#qq plot - normality
qqnorm(toxic.fit$residuals, ylab="Residuals")
qqline(toxic.fit$residuals, col="red")
```
    
\newpage    
    
### (c)

```{r eval=FALSE, warning=FALSE, include=FALSE,echo=F}
#c
#(ii)
#new model with scaled log2x
toxic.r <- toxic
toxic.r$log2x <- toxic.r$log2x - 4
toxic.fit.r <- glm(mortality ~ 0+log2x, family = quasibinomial(link = "logit"), data = toxic.r)
toxic.sum.r<-summary(toxic.fit.r)

#likelihood test
lr.stat <- toxic.fit.r$deviance - toxic.fit$deviance
pvalue.H0.lr <- pchisq(q = lr.stat, df = 1, lower.tail = F)
```
   
#### (iii)

I didn't understand what "plot the restricted log likelihood against the hypothesized value of log2(gamma)" is, because restricted log likelihood is one number, and the hypothesized value of log2(gamma) is 4 (null hypothesis in the previous part), so if I were to plot it it would be just a single point in the plane, so I did not understand how that would help with constructing a confidence interval.    
  I used some built in function to generate the 95% confidence interval for the parameter of the restricted model to be (0.083, 1.14), so the 95% confidence interval for gamma would be (1.06, 2.20). This also doesn't really make sense because we fail to reject that gamma = 4 earlier, but 4 is outside of this confidence interval.
```{r echo=FALSE, warning=FALSE,eval=F}
#construct likelihood based confidence set for gamma
2**confint(toxic.fit.r,level = 0.95)

```



```{r echo=FALSE, warning=FALSE,eval=F}
#(d) (v)
fit.pearson.0 <- glm(Y ~ color, family = binomial(link = "logit"), data = Data)
toxic.fit <- glm(mortality ~ log2x, family = quasibinomial(link = "logit"), data = toxic)

Observed <- t(sapply(1:length(unique(toxic$log2x)), function(i){ind.i <- toxic$log2x==i; c(sum(toxic$mortality[ind.i]),sum(ind.i)-sum(toxic$mortality[ind.i]))}))
rownames(Observed) <- c("1", "2", "3", "4","5","6"); colnames(Observed) <- c("Y=1", "Y=0")

predicted <- exp(predict(toxic.fit))/(1+exp(predict(toxic.fit)))
Expected <- t(sapply(1:length(unique(toxic$log2x)), function(i){ind.i <- toxic$log2x==i; c(sum(predicted[ind.i]),sum(ind.i)-sum(predicted[ind.i]))}))
rownames(Expected) <- c("1", "2", "3", "4","5","6"); colnames(Expected) <- c("Y=1", "Y=0")

Pearson.X2 <- sum((Observed-Expected)^2/Expected, na.rm = T)
pchisq(q = Pearson.X2, df = 6-2, lower.tail = F)
```

