---
title: "HW8"
author: "Giang Vu"
date: "4/3/2021"
output: pdf_document
---

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
library(dplyr)
library(plyr)
library(tidyr)
library(ggplot2)
library(purrr)
library(tibble)
#setwd("/Users/giangvu/Desktop/STAT 2132 - Applied Stat Method 2/HW/HW9")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,warning=FALSE,out.width="80%")
```


## 4.\

### (a)\

The first model's estimate for $\sigma$ is 11.6, and the second model's estimate is 4.3.\

The second model with only treatments C and D has smaller estimate, probably because C and D are of the same drug type so the variance among them is smaller, unlike when we include A and B which are completely different treatments.\

```{r}
#load data
sleep <- data.frame(read.table(file = "sleep.txt", header = T, sep = "\t", stringsAsFactors = F))
sleep$Individual <- as.factor(sleep$Individual)
sleep$Treatment <- as.factor(sleep$Treatment)

#additive fixed effect model
sleep.fit1 <- aov(Hours ~ Individual + Treatment, data = sleep)
#estimate sigma by sqrt(MSE)
summary(sleep.fit1) #trt effect is not signi
sqrt(134.3) #model 1 estimate

#another additive but w/o B and A
sleep.fit2 <- aov(Hours ~ Individual + Treatment, data = sleep[sleep$Treatment %in% c("C","D"),])
#estimate sigma by sqrt(MSE)
summary(sleep.fit2)
sqrt(18.3) #model 2 estimate

#second model with only C and D has smaller estimate, probably because C and D are of the same drug type so the variance is smaller, unlike A and B which are completely different treatment.
```

### (b)\

Because the mean squared error for the model 1 (full data) is much larger than model 2 (only treatments B and C), and also the estimate for the treatment effect in model 2 is not statistically significant, it would be reasonable to expect that model 1 fails to account for some of the treatment effect difference between A, B, and C and D. Therefore, we have some evidence to believe that adding an interaction term in our model would be useful here to eliminate this issue with problem 1.

\newpage

